Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex

\pointRaised{R1}{%
	Section 3, full model: The square brackets seem to be misplaced, as the exponent includes indices of the sum contained in the brackets.
}

\reply{%
	\textbf{The expression is now corrected.}
}

\pointRaised{R1}{%
Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in that they seem to mix the posterior behavior (e.g., parameters are updated though standard multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…). This subsection might be easier to understand if its purpose was more bluntly described in an opening sentence or paragraph and streamlined to focus only on the full conditional distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in intermediate conclusions. In particular, the derivation of the probability of $\Gamma_{.j}$ seems to be missing a factor of $u_{fl}^I(gamma_{ij}^f=l)$. This is not important to the eventual full conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
}

\reply{%
	\textcolor{red}{We have revised our presentation of the Gibbs sampler with this feedback in mind. The main body of the text now presents the final full conditionals, with revised derivations moved now in the appendix. In particular, we have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented.}
}

%\reply{%
%\textcolor{red}{\todo We should meet specifically to talk about how to best present the Gibbs Sampler. I think it be good to move as much of it as possible (at least the messy math and derivations) into the appendix.} \textcolor{blue}{I agree that these should be moved to the appendix as well.}
%
%We have included an extra step in the derivation of $p(\Gamma_{.j})$ that shows exactly how we divide by $\prod_{i = 1}^{n_1}\prod_{f = 1}^{F} \prod_{\ell = 1}^{L_f}  u_{fl}^{I(\gamma_{ij}^f=l)}$ and use proportionality to arrive that the distribution presented. \textcolor{red}{I want you guys to look at it though. Its pretty ugly.}
%}

@@

\pointRaised{R1}{%
Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the conditional distribution of Z was included using the summation-style notation rather than the binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
presented on the first line of page 8.
}

\reply{%
	\textbf{The contribution of $Z$ in the full conditional is just the prior distribution, which we have expressed with summation notation to reflect the piecewise nature of the prior. The binomial-looking expression at the top of page 8 is the posterior distribution for $\pi$. We note that the posterior $p(\pi| \Gamma, Z, \Phi)$ in fact only depends on $Z$, and is thus shortened to $p(\pi|Z)$. We have updated our explanation to make this more clear.}
}

@@

\pointRaised{R1}{%
Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
}

\reply{%
	\textbf{We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.}
}

\pointRaised{R1}{%
Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It would be helpful to provide a brief justification of integrating out pi.
}
\reply{%
	\textcolor{red}{I don't fully understand the first part of the comment} \textcolor{blue}{I believe the person is asking you to make a connection to your full conditionals to how you make Gibbs draws. To do this, state that now you have all the full conditionals to run a Gibbs sampler, which can be found in Appendix XX.}
	We have included the details for integrating out $\pi$.
}

\pointRaised{R1}{%
Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. (1) My understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to (should) be identified in the posterior distribution. (2) Thus, you take a small subsample of the records of size $S < H_{j_p}$, and store these in $R^{SEI}$ rather than the complete index of pairs in each R. (3) Can you quantify the computational savings due to this method? (4) Do you have advice for how small S can be? (5) In your simulations, how does the use of this method or choice of S affect the results?
}

\reply{%
\textcolor{blue}{Brian: There are five questions here that should be answered. The first two should be yes, the understanding is correct or no, it's not correct. (If the reviewer doesn't understand, then provide an explanation of your approach. For (3), we quantify the SEI procedure in Section 4.2, where we state: 
"Rather than storing \(n_1 \times n_2\) record labels, SEI allows us to store at most \(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be." SEI does not affect the computation time of the Gibbs sampler, since the complexity of the sampler is $O(n_2 P)$, regardless of how many comparison vectors are of pattern $P$. Turning to (4), in practice, we recommend $S=10$, as this reduces the number of stored indices for highly unlikely record pairings, but does is not likely to eliminate any of the indices for likely matches. Choosing $S$ too low can concentrate undue mass on unlikely matches and distort linkage results. I don't see an answer to the question (5).}
}

\pointRaised{R1}{%
Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
}

@@

\pointRaised{R1}{%
	Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results hold over multiple simulated datasets, Or was this just a chance result for each setting?
}

\reply{%
	In the caption under Figure 3, we explain that we have 100 pairs of sets of records for each level of overlap and error. Thus the table represents results over 900 pairs of datasets. We have added this description in the body of the paper as well for clarity. 
}

\pointRaised{R1}{%
Section 5.3 simulation – RR results are not presented.
}

\reply{%
	\textcolor{red}{We have updated the \textcolor{blue}{Figure XXXX} with these results. Since its is presented along side NPV and PPV, for which values close to 1 indicate strong performance, we have presented decision rate (DR) results.}
}

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex

\pointRaised{R1}{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
}
\reply{%
	\textbf{This has been corrected.}
}

\pointRaised{R1}{%
Section 3.1, pmf for $Gamma_{.j}$ and full conditional for Z: The notation in the final equation may be more clear if you replace i with $z_j$, as in $w_{z_j j}$. I believe that “$n_1+1$” should be “$n_1+ j$” in several places.
}

\reply{%
	\textbf{We have changed $w_{ij}$ to $w_{z_j, j}$ in the pmf for $\Gamma_{.j}$ and the full conditional for $Z_j$. We have been sure that “$n_1+1$” has been changed to “$n_1+ j$” throughout the paper.}
}

\pointRaised{R1}{%
Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It would be helpful to provide a brief justification of integrating out pi.
}
\reply{%
 \textbf{We have revised the presentation of the Gibbs sampler based on this feedback. The main body of the paper presents the results, and revised derivations (including justification for integrating out $\pi$) have been placed in the appendix.}
}

### 2022-09-09 ba_resubmit/response_to_reviewers_rcs.tex

\pointRaised{R1}{%
Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
}

\reply{%
	\textbf{The lemma now includes assumptions/definitions for all variables used within the statement.}
}

\pointRaised{R1}{%
Section 5.1 simulations – there is a detailed comparison for one set of simulation settings, including a binary definition of a “match”. Do these results differ as these settings change?
}

\reply{%
	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl. The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity. \textcolor{red}{To illustrate this, we have included in the appendix an additional simulation study under different comparison vector settings.}
}

%\reply{%
%	Regardless of the specific construction of the comparison vectors, the computational complexity the Gibbs sampler under BRL is $O(n_1 n_2)$, where as it is $O(n_2 P)$ for fabl. The exact shape of the lines will be different, but in all cases, fabl removes dependence on $n_1$ in the computational complexity. 
%	
%	\textcolor{red}{I think this is clear in the paper, so I did not add anything to the actual text.} \textcolor{blue}{I'm not sure that you answered the reviewers question. You show a comparison for one set of simulation settings, but if you change the settings what would occur? I think the reviewer is asking for more clarify regarding how the lines would change. You may wish to give a few different scenarios to illustrate how things change?}
%}

### 2022-09-09 ba_resubmit/deprecated/response_markdown.Rmd

**For example, $I_obs()$ is never defined, $n_12(Z)$ first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.**
\begin{quote}
\textcolor{red}{How and when should I define $I_{obs}$? I probably need to lay out certain assumptions about missingness?}

Thank you for identifying inconsistencies with the notation $n_{12}(Z)$. We have made these corrections.
\end{quote}

**Throughout the manuscript, $n_1$ and $n_2$ are used interchangeably with $n_A$ and $n_B$, sometimes in the same section, as it happens, for example, in Section 3.1.**
\begin{quote}
Thank you for addressing this oversight. We have made these corrections throughout the text.
\end{quote}

**The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein $(\theta_{11},$ $if Z_j, \hat{Z_j}, …)$.**
\begin{quote}
Thank you for addressing this oversight. The previous expression $Z_j, \hat{Z}_j \leq n_1$ was meant to be understood as $Z_j\leq n_1, \hat{Z}_j \leq n_1$ We agree that the latter expression is more clear, and have made that change.
\end{quote}

@@

**In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)**
\begin{quote}
We have updated the Lemma to include the definition of $B$.
\end{quote}

**Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.  For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.**
\begin{quote}
\textcolor{red}{Is this true? I don't mind making the change, I've just never seen commas in that way.}
\end{quote}

@@

# Reviewer 3

The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation results are presented and the variant is applied on a new, significantly larger data set.

A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in section 4. It is simple but consequential. The simulations and applications show how the variant accelerates the computation relative to the original method and reveals the trade-offs. The value of the paper resides in that computational complexity continues to be a difficult obstacle in the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s approach to an extent not possible before.

**It is important to present the research for what it is: An extension of the method of Sadinle motivated by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should be made clearer, beginning in the abstract and reiterated in the conclusion.**
\begin{quote}
\textcolor{red}{I'm not sure how to address this.}
\end{quote}

### 2022-09-09 ba_resubmit/deprecated/kundinger-submit-final.tex
clean up.

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

\begin{align*}
	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
\end{align*}
Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.

>>>>

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12} = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

\begin{align*}
	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
\end{align*}
Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.
